---
title: "CTSL production experiment"
output: rmarkdown::github_document
#output:
  #html_document: default
  #pdf_document: default
---

```{r import, include=FALSE}
#this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
#setwd(this.dir)

library("tidyverse")
library("xlsx")
library("readxl")
source("helpers.R")
library("wesanderson")
palette = wes_palette("Zissou1")
theme_set(theme_bw())

files<-list.files(path="../data", pattern=".xlsx", full.names = TRUE)

dd = NULL

for (f in files) {
  print(f)
  d = read_excel(f)
  #print(unique(d$speakerName))
  dd = rbind(dd,d)
}

dd = dd %>% drop_na(speakerName)

unique(dd$speakerName)
```

**Number of participants, number of trials each participant completed:**
```{r number_tested, echo=FALSE}
length(unique(dd$speakerName)) # 15 people

# this shouldn't be done manually -- extract this from original files or create new file with demographic data and load it here
dd = dd %>% mutate(language = ifelse(speakerName=="aysetas","Turkish",ifelse(speakerName=="hatice","Turkish",ifelse(speakerName=="aliamca","Turkish",ifelse(speakerName=="durdukılıç(türkçe)","Turkish",ifelse(speakerName=="Bülent(türkçe)", "Turkish","CTSL"))))))

table(dd$speakerName,dd$language) # 5 Turkish, 6 CTSL (all completed 60 trials)
```

**Overall accuracy:**
```{r ovv_accuracy, echo=FALSE}
toplot =  dd %>%
  group_by(speakerName,language) %>%
  dplyr::summarise(Mean=mean(correct),CILow=ci.low(correct),CIHigh=ci.high(correct)) %>%
  ungroup() %>%
  mutate(YMin=Mean-CILow,YMax=Mean+CIHigh) %>%
  mutate(lowacc=ifelse(Mean<0.70,"1","0"))

h=0.70
ggplot(toplot, aes(x=reorder(speakerName,Mean), y=Mean, fill=language)) +
  geom_bar(stat="identity") +
  geom_hline(yintercept=h) +
  geom_text(aes(0, h, label=h, vjust=-1, hjust=-0.3)) +
  geom_errorbar(aes(ymin = YMin, ymax = YMax),width=.25) +
  theme(axis.text.x=element_text(angle=45, hjust=1),axis.title.x=element_blank()) +
  ylab("Accuracy") +
  scale_fill_manual(values=c("darkgreen","darkgoldenrod1"))

ggsave(file="../graphs/accuracy.pdf",width=6,height=4.5)

# to exclude the participants with low(?) accuracy
#exclude = toplot %>%
#  filter(lowacc==1)
#exclude

#dd = dd[!(dd$speakerName %in% exclude$speakerName),]
```

**Accuracy by trial type:**
```{r accuracy_tt, echo=FALSE}
toplot =  dd %>%
  group_by(language,trialType) %>%
  dplyr::summarise(Mean=mean(correct),CILow=ci.low(correct),CIHigh=ci.high(correct)) %>%
  ungroup() %>%
  mutate(YMin=Mean-CILow,YMax=Mean+CIHigh) %>%
  mutate(lowacc=ifelse(Mean<0.70,"1","0"))

ggplot(toplot, aes(x=reorder(trialType,Mean), y=Mean, fill=language)) +
  geom_bar(position="dodge", stat="identity") +
  geom_errorbar(aes(ymin = YMin, ymax = YMax),width=.25, position=position_dodge(width = 0.9)) +
  theme(axis.title.x=element_blank()) +
  ylab("Accuracy") +
  scale_fill_manual(values=c("darkgreen","darkgoldenrod1"))

ggsave(file="../graphs/accuracy_trialType.pdf",width=6,height=4.5)

```

**Overinformativeness trials:**
```{r overinformativeness, echo=FALSE}
df = dd %>%
  filter(trialType == "overinformativeness")

#nrow(df) # 330 overinformativeness trials
drows = nrow(df)
#table(df$condition)

df = df %>%
  mutate(targetFull=targetName) %>%
  separate(targetFull,c("targetobj","targetsize","targetcolor"))

table(df$targetobj,df$correct)
```

**Was a color mentioned?**
```{r colormention, echo=FALSE}
colors = "blue|green|red|brown|black|clear|purple|pink|silver|gray|grey|white|yellow|violet|gold|silver|orange"
df$colorMention = ifelse(grepl(colors, df$gloss, ignore.case = TRUE), 1,0)
table(df$colorMention)

# checking manual coding for color
#df$colorManual = ifelse(grepl("color", df$pos_sequence, ignore.case = TRUE), 1,0)
#table(df$colorManual)
```

**Was size mentioned?**
```{r sizemention, echo=FALSE}
sizes = "big|small|biggest|smalles|tiny|huge|little"
df$sizeMention = ifelse(grepl(sizes, df$gloss, ignore.case = TRUE), 1,0)
table(df$sizeMention)

# checking manual coding for size
#df$sizeManual = ifelse(grepl("size", df$pos_sequence, ignore.case = TRUE), 1,0)
#table(df$sizeManual)
```

**Was both a color and size mentioned?**
```{r colorsizesmention, echo=FALSE}
df$colorsizeMention = ifelse(df$colorMention==1&df$sizeMention==1,1,0)
table(df$colorsizeMention)
```

**Was an object name mentioned?**
```{r objectmention, echo=FALSE}
names  = "dress|apple|hair|pepper|glasses|ball|lighter|pot|button|clock|knife|tie|umbrella|spoon|shorts|tshirt|brush|box|skirt|socks|scissors|glove|lock|shoe|flower|bag|lipstick|toothbrush|car|baloon|balloon|chair|key|pants|";
#df$objectMention = ifelse(grepl(names, df$gloss, ignore.case = TRUE), 1, 0)
#table(df$objectMention)

# checking manual coding for noun
df$objectManual = ifelse(grepl("noun", df$pos_sequence, ignore.case = TRUE), 1,0)
table(df$objectManual)
```

**In how many trials did the listener choose the right object?**
```{r fcorrect_selection, echo=FALSE}
table(df$correct)
non_target_click = drows-sum(df$correct)

print(paste("percentage of trials where target wasn't selected: ", non_target_click*100/drows)) #8.89
print("distribution accross conditions: ")
table(df$condition,df$correct)
```

**Exclude trials where target wasn't selected**

```{r clicktarget, echo=FALSE}
targets = df %>% filter(correct == "1")
nrow(targets) # 411 trials

targets = targets %>%
  mutate(UtteranceType = ifelse(colorMention==1 & sizeMention==1,"color and size",ifelse(colorMention==1,"only color",ifelse(sizeMention==1,"only size","other")))) %>%
  filter(UtteranceType!="other") %>%
  mutate(rowNum = row_number()) %>%
  mutate(onlyColorMention = ifelse(UtteranceType =="only color",1,0)) %>%
  mutate(onlySizeMention = ifelse(UtteranceType =="only size",1,0)) %>%
  mutate(RedundantProperty = ifelse(condition=="color_sufficient","size", ifelse(condition=="size_sufficient","color", NA))) %>%
  mutate(SufficientProperty = ifelse(condition=="color_sufficient","color", ifelse(condition=="size_sufficient","size", NA))) %>%
  mutate(RedExp3Cat = ifelse(UtteranceType=="color and size","redundant",ifelse((SufficientProperty=="color"&UtteranceType=="only color"),"minimal",ifelse((SufficientProperty=="size"&UtteranceType=="only size"),"minimal","underinformative")))) %>%
  mutate(RedExp2Cat = ifelse(RedExp3Cat=="redundant","redundant","not_redundant")) %>%
  mutate(redundant = ifelse(RedExp2Cat=="redundant",1,0)) %>%
  mutate(condition=ifelse(condition=="size_sufficient","size sufficient",ifelse(condition=="color_sufficient","color sufficient",NA))) %>%
  select(gameid,speakerName,language,roundNum,condition,targetName,correct,gloss,colorMention,sizeMention,UtteranceType,onlyColorMention,onlySizeMention,colorsizeMention,RedundantProperty,SufficientProperty,RedExp3Cat,RedExp2Cat,redundant)
```

**Proportion of utterance by utterance type and trial type**
```{r mainplot, echo=FALSE}
agr = targets %>%
  gather(MentionType,mention,onlyColorMention:colorsizeMention) %>%
  group_by(RedundantProperty,condition,MentionType) %>%
  summarise(Mean=mean(mention),CIlow=ci.low(mention),CIhigh=ci.high(mention)) %>%
  ungroup() %>%
  mutate(Ymin=Mean-CIlow, Ymax=Mean+CIhigh) %>%
  mutate(MentionType=fct_recode(MentionType,"color and size"="colorsizeMention","only color"="onlyColorMention","only size"="onlySizeMention"))

ggplot(agr, aes(x=condition,y=Mean,fill=RedundantProperty)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = Ymin, ymax = Ymax),width=.25) +
  xlab("Trial type") +
  facet_grid(~MentionType) + 
  scale_fill_manual(values=c(palette[1],palette[3])) +
  theme(legend.position = "none",axis.title.x = element_blank(),axis.text.x = element_text(size=7)) +
  ylab("Proportion of utterance")

ggsave(file="../graphs/proportion_overall.pdf",width=6,height=3)
```

**Proportion of utterance by language, utterance type and trial type**
```{r mainplot_language, echo=FALSE}
agr = targets %>%
  gather(MentionType,mention,onlyColorMention:colorsizeMention) %>%
  group_by(RedundantProperty,condition,MentionType,language) %>%
  summarise(Mean=mean(mention),CIlow=ci.low(mention),CIhigh=ci.high(mention)) %>%
  ungroup() %>%
  mutate(Ymin=Mean-CIlow, Ymax=Mean+CIhigh) %>%
  mutate(MentionType=fct_recode(MentionType,"color and size"="colorsizeMention","only color"="onlyColorMention","only size"="onlySizeMention"))
  
ggplot(agr, aes(x=condition,y=Mean,fill=RedundantProperty)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = Ymin, ymax = Ymax),width=.25) +
  xlab("Trial type") +
  facet_grid(language ~ MentionType) + 
  scale_fill_manual(values=c(palette[1],palette[3])) +
  theme(legend.position = "none",axis.title.x = element_blank(),axis.text.x = element_text(size=7))+
  ylab("Proportion of utterance")

ggsave(file="../graphs/proportion_language.png",width=6,height=4)
```

```{r simple_model, echo=FALSE}
library(lme4)
library(languageR)
#centering condition 
tomodel = targets %>%
  mutate(condition = as.factor(condition)) %>%
  mutate(ccondition = as.numeric(condition) - mean(as.numeric(condition)))

m = glmer(redundant ~ ccondition + (1+ccondition|speakerName) + (1|targetName), data=tomodel, family="binomial")
summary(m)
```
**Proportion of utterance type**
```{r utterancetype, echo=FALSE}
agr = targets %>%
  mutate(ThreeRedundant=ifelse(RedExp3Cat=="redundant",1,0)) %>%
  mutate(ThreeMinimal=ifelse(RedExp3Cat=="minimal",1,0)) %>%
  mutate(ThreeUnderinformative=ifelse(RedExp3Cat=="underinformative",1,0)) %>%
  gather(ThreeType,value,ThreeRedundant:ThreeUnderinformative) %>%
  group_by(RedundantProperty,ThreeType,condition,language) %>%
  summarise(Mean=mean(value),CIlow=ci.low(value),CIhigh=ci.high(value)) %>%
  ungroup() %>%
  mutate(Ymin=Mean-CIlow, Ymax=Mean+CIhigh) %>%
  mutate(ThreeType=fct_recode(ThreeType,"redundant"="ThreeRedundant","underinformative"="ThreeUnderinformative","minimal"="ThreeMinimal"))

ggplot(agr, aes(x=condition,y=Mean,fill=RedundantProperty)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = Ymin, ymax = Ymax),width=.25) +
  xlab("Trial type") +
  facet_grid(language ~ ThreeType) + 
  scale_fill_manual(values=c(palette[1],palette[3])) +
  theme(legend.position = "none",axis.title.x = element_blank(),axis.text.x = element_text(size=7))+
  ylab("Proportion of utterance")

ggsave(file="../graphs/proportion_threetype.png",width=6,height=4)
``` 

**Proportion of utterance by utterance type and trial type - for each participant**    
see graphs/by_participant.pdf
```{r mainplot_participant, include=FALSE}
agr = targets %>%
  gather(MentionType,mention,onlyColorMention:colorsizeMention) %>%
  group_by(RedundantProperty,condition,MentionType,speakerName) %>%
  summarise(Mean=mean(mention),CIlow=ci.low(mention),CIhigh=ci.high(mention)) %>%
  ungroup() %>%
  mutate(Ymin=Mean-CIlow, Ymax=Mean+CIhigh) %>%
  mutate(MentionType=fct_recode(MentionType,"color and size"="colorsizeMention","only color"="onlyColorMention","only size"="onlySizeMention"))

ggplot(agr, aes(x=condition,y=Mean,fill=condition)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = Ymin, ymax = Ymax),width=.25) +
  xlab("Trial type") +
  facet_wrap(speakerName ~ MentionType, nrow=16) + 
  #scale_fill_discrete(guide = FALSE) +
  scale_fill_manual(values=c(palette[1],palette[3])) +
  theme(axis.title.x = element_blank(),axis.text.x = element_text(size=7))+
  ylab("Proportion of utterance") +
  scale_alpha_manual(values=c(0.5,1))

ggsave(file="../graphs/by_participant.png",width=6,height=30)
```

**Order of constituents**    
see graphs/all_constituents.pdf
```{r order, include=FALSE}
df_long = df %>%
  mutate(pos_sequence2=pos_sequence) %>%
  separate(pos_sequence2,c("attempt1","attempt2","attempt3","attempt4","attempt5"), sep=",") %>%
  gather(attempt,pos,attempt1:attempt5) %>%
  drop_na(pos) %>%
  mutate(pos = trimws(pos,which=c("left")))

remove = "\\?"

toplot = df_long %>%
  filter(!str_detect(pos,remove)) %>%
  group_by(pos,language) %>%
  count()

ggplot(toplot, aes(x=reorder(pos,n),y=n,fill=language)) +
  geom_bar(position="dodge", stat = "identity") +
  coord_flip() +
  scale_fill_manual(values=c("darkgreen","darkgoldenrod1"))

ggsave(file="../graphs/all_constituents.pdf",width=8,height=22)
```

**Order of unique constituents (only include first occurance of each constituent)**   
see graphs/unique_constituents.pdf
```{r order_2, include=FALSE}
remove = "\\?"

toplot = df_long %>%
  filter(!str_detect(pos,remove)) %>%
  mutate(itemized = strsplit(pos," "))

toplot$uniqueItemized=c()
toplot$uniquePos=c()
for (i in 1:length(toplot$itemized)){
  toplot$uniqueItemized[i]=map(toplot$itemized[i], ~ unique(.x))
  toplot$uniquePos[i] = map(toplot$uniqueItemized[i], ~ paste(.x,collapse=" "))
}

toplot = toplot %>%
  mutate(uniquePos = as.character(uniquePos)) %>%
  mutate(uniquePos = trimws(uniquePos,which=c("right"))) %>%
  group_by(uniquePos,language) %>%
  count()

ggplot(toplot, aes(x=reorder(uniquePos,n),y=n,fill=language)) +
  geom_bar(position="dodge", stat = "identity") +
  coord_flip() +
  scale_fill_manual(values=c("darkgreen","darkgoldenrod1"))

ggsave(file="../graphs/unique_constituents.pdf",width=8,height=22)
```

**Order of unique color, noun, size (remove all other constituents)**
```{r order_3, echo=FALSE}
replace = "noun2|color2|size2|mod2|mod3|verb2|num2|clr|neg|loc|other|verb|num|OTHER|COLOR|\\?|QUESTION|mod|clor|color-size|noun-size|lr|nnoun|sizeB|noun4|mod4|noun-|cerb2"

toplot = df_long
toplot$replacedPos = gsub("  ", " ", toplot$pos)
toplot$replacedPos = gsub(replace, "", toplot$replacedPos)
toplot$replacedPos = gsub("  ", " ", toplot$replacedPos)
toplot$replacedPos = gsub("  ", " ", toplot$replacedPos)

toplot = toplot %>%
  mutate(replacedPos = trimws(replacedPos,which=c("left"))) %>%
  mutate(itemized = strsplit(replacedPos," "))

toplot$uniqueItemized=c()
toplot$uniquePos=c()
for (i in 1:length(toplot$itemized)){
  toplot$uniqueItemized[i]=map(toplot$itemized[i], ~ unique(.x))
  toplot$uniquePos[i] = map(toplot$uniqueItemized[i], ~ paste(.x,collapse=" "))
}

toplot = toplot %>%
  mutate(uniquePos = as.character(uniquePos)) %>%
  mutate(uniquePos = trimws(uniquePos,which=c("right"))) %>%
  group_by(uniquePos,language) %>%
  count()

toplot <- toplot[-which(toplot$uniquePos == ""), ]

ctsl = toplot %>%
  filter(language=="CTSL")

ggplot(toplot, aes(x=reorder(uniquePos,n),y=n,fill=language)) +
  geom_bar(position="dodge", stat = "identity") +
  coord_flip() +
  scale_fill_manual(values=c("darkgreen","darkgoldenrod1")) +
  theme(axis.title.y=element_blank()) +
  ylab("count")

ggsave(file="../graphs/unique_only_constituents.pdf",width=8,height=5)
```

**Was negation used**   
Per attempt (not per trial)
```{r negation, echo=FALSE}
df_long$negation = ifelse(grepl("neg", df_long$pos, ignore.case = TRUE), 1,0)

table(df_long$negation)
```

**Word before negation & word after negation**   
```{r negation2, echo=FALSE}
negation = df_long %>%
  filter(negation=="1") %>%
  select(speakerName,gloss,pos,language) %>%
  mutate(before = ifelse(startsWith(pos,"neg"),"",sub(" neg.*","",pos))) %>%
  #mutate(before = sub(" neg.*","",pos)) %>%
  mutate(wordBefore = word(before,-1)) %>%
  mutate(after = ifelse(endsWith(pos,"neg"),"",sub(".*neg ","",pos))) %>%
  #mutate(after = sub(".*neg ","",pos)) %>%
  mutate(wordAfter = word(after,1)) %>%
  group_by(wordBefore,wordAfter,language) %>%
  count() %>%
  mutate(negationString=paste(wordBefore,"neg",wordAfter)) %>%
  mutate(negationString = trimws(negationString,which=c("right"))) %>%
  mutate(negationString = trimws(negationString,which=c("left")))
  
 ggplot(negation, aes(x=reorder(negationString,n),y=n,fill=language)) +
  geom_bar(position="dodge", stat = "identity") +
  coord_flip() +
  scale_fill_manual(values=c("darkgreen","darkgoldenrod1")) +
  theme(axis.title.y=element_blank())
 
ggsave(file="../graphs/negation.png")

```

**Adjective ordering trials:**
```{r adjordering, echo=FALSE}
da = dd %>%
  filter(trialType == "adjective_ordering") %>%
  mutate(targetName2=targetName) %>%
  separate(targetName2,c("targetobj","targetadj1","targetadj2"))

table(da$targetobj,da$correct)

toplot =  da %>%
  group_by(targetobj) %>%
  dplyr::summarise(Mean=mean(correct),CILow=ci.low(correct),CIHigh=ci.high(correct)) %>%
  ungroup() %>%
  mutate(YMin=Mean-CILow,YMax=Mean+CIHigh)

ggplot(toplot, aes(x=reorder(targetobj,Mean), y=Mean)) +
  geom_bar(position="dodge", stat="identity", fill="lightblue") +
  geom_errorbar(aes(ymin = YMin, ymax = YMax),width=.25, position=position_dodge(width = 0.9)) +
  theme(axis.title.x=element_blank()) +
  ylab("Accuracy") +
  scale_fill_manual(values=c("darkgreen","darkgoldenrod1")) +
  theme(axis.title.y=element_blank())

ggsave(file="../graphs/accuracy_adjordering.pdf",width=8,height=4)
```

**Best objects: frog, hair**  
man2 not included (need to fix typos!!)
```{r comment='', echo=FALSE, results='asis'}
da_long = da %>%
  mutate(gloss_long=gloss) %>%
  separate(gloss,c("attempt1","attempt2","attempt3","attempt4","attempt5"), sep=",") %>%
  gather(attempt,gloss,attempt1:attempt5) %>%
  drop_na(gloss) %>%
  mutate(gloss = trimws(gloss,which=c("left")))

frog = da_long %>%
  filter(targetobj=="frog") %>%
  select(language,targetName,gloss)
print("FROG")
knitr::kable(frog, floating.environment="sidewaystable")

hair = da_long %>%
  filter(targetobj=="hair") %>%
  select(language,targetName,gloss)
print("HAIR")
knitr::kable(hair, floating.environment="sidewaystable")
```

**Worst objects: woman2, nails**   
```{r comment='', echo=FALSE, results='asis'}
woman2 = da_long %>%
  filter(targetobj=="woman2") %>%
  select(language,targetName,gloss)
print("WOMAN2")
knitr::kable(woman2, floating.environment="sidewaystable")

nails = da_long %>%
  filter(targetobj=="nails") %>%
  select(language,targetName,gloss)
print("NAILS")
knitr::kable(nails, floating.environment="sidewaystable")
```

**Next**    
- Get cohort information from Rabia (facet all plots by cohort/age)    
- Fix all typos (COLOR to clr, fix ?)    
- Coding pos for adjective ordering trials  
- Coding order of adjectives
