---
title: "English production experiment"
output: rmarkdown::github_document
#output:
  #html_document: default
  #pdf_document: default
---

```{r import, include=FALSE}
#this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
#setwd(this.dir)

library("tidyverse")
library("xlsx")
library("readxl")
source("helpers.R")
library("wesanderson")
palette = wes_palette("Zissou1")
theme_set(theme_bw())

demo = read.csv(file="../data/example-subject_information.csv")
dd = read.csv(file="../data/merged_data.csv",sep="\t", header=T, quote="")
```

```{r data_clean, include=FALSE}
dd = dd %>%
  mutate(targetName = ifelse(clickedObjTargetStatus=="target",as.character(clickedObjName),ifelse(alt1TargetStatus=="target",as.character(alt1Name),ifelse(alt2TargetStatus=="target",as.character(alt2Name),ifelse(alt3TargetStatus=="target",as.character(alt3Name),NA))))) %>%
  mutate(trialType = ifelse(clickedObjTrialType=="overinformativeness", "redundantModification", ifelse(clickedObjTrialType=="adjective_ordering","modifierOrdering", as.character(clickedObjTrialType)))) %>%
  mutate(correct = ifelse(clickedObjTargetStatus=="target",1,0))

demo = demo %>%
  mutate(role = ifelse(role=="speaker", "director", ifelse(role=="listener", "guesser", NA)))
         
  #mutate(targetlabel = targetname) %>%
  #separate(targetlabel, c("targetobject","targetmaterial","targetcolor","targetoriginal"), sep="_") %>%
  #mutate(clickedObjlabel = clickedObjName) %>%
  #separate(clickedObjlabel, c("clickedobject","clickedmaterial","clickedcolor","clickedoriginal"), sep = "_")

dd$extra = NULL
```

**Number of participants tested:**
```{r number_tested, echo=FALSE}
length(demo$workerid)
table(demo$role)
```

**Number of unique gameids and number of rounds for each game**
```{r roundno, echo=FALSE}
length(unique(dd$gameid))

round_no = dd %>%
  group_by(gameid) %>%
  summarize(total_round_number = n())

table(round_no$total_round_number)
```


```{r df_comparison, eval=FALSE, include=FALSE}
# **Number of gameids in the server vs mturk data:**

#TODO: fix this using unique.R
# serverGameids =  df %>% select(gameid) %>% unique()
# length(serverGameids$gameid)
# 
# mturkGameids = demo %>%select(gameID) %>% group_by(gameID) %>% count()
# #length(mturkGameids$gameID)
# length(unique(mturkGameids$gameID))
```

**Exclusion: Remove games with non-native speakers: **

```{r exclusions, echo=FALSE}
notnative = demo %>% filter(demo$nativeEnglish != "yes")  %>% select(gameID) #7754-6
notnative

dd = dd[!(dd$gameid %in% notnative$gameID),]
```

**Overall accuracy:**
```{r ovv_accuracy, echo=FALSE}
toplot =  dd %>%
  filter(trialType=="redundantModification") %>%
  group_by(gameid) %>%
  dplyr::summarise(Mean=mean(correct),CILow=ci.low(correct),CIHigh=ci.high(correct)) %>%
  ungroup() %>%
  mutate(YMin=Mean-CILow,YMax=Mean+CIHigh) %>%
  mutate(lowacc=ifelse(Mean<0.70,"1","0"))

h=0.70
ggplot(toplot, aes(x=reorder(gameid,Mean), y=Mean)) +
  geom_bar(stat="identity", fill="lightblue") +
  geom_hline(yintercept=h) +
  geom_text(aes(0, h, label=h, vjust=-1, hjust=-0.3)) +
  geom_errorbar(aes(ymin = YMin, ymax = YMax),width=.25) +
  theme(axis.text.x=element_blank()) +
  ylab("Accuracy") +
  xlab("English speakers")

ggsave(file="../graphs/accuracy_redundant.pdf",width=4,height=2.5)
```

**Exclusion: Remove games with accuracy<70%: **

```{r exc_accuracy, echo=FALSE}
exclude = toplot %>%
  filter(lowacc==1)
exclude # 10 games

dd = dd[!(dd$gameid %in% exclude$gameid),]

length(unique(dd$gameid)) #46 games left
```

**Accuracy by trial type:**
```{r accuracy_tt, echo=FALSE}
toplot =  dd %>%
  group_by(trialType) %>%
  dplyr::summarise(Mean=mean(correct),CILow=ci.low(correct),CIHigh=ci.high(correct)) %>%
  ungroup() %>%
  mutate(YMin=Mean-CILow,YMax=Mean+CIHigh) %>%
  mutate(lowacc=ifelse(Mean<0.70,"1","0"))

ggplot(toplot, aes(x=reorder(trialType,Mean), y=Mean)) +
  geom_bar(position="dodge", stat="identity", fill="lightblue") +
  geom_errorbar(aes(ymin = YMin, ymax = YMax),width=.25, position=position_dodge(width = 0.9)) +
  theme(axis.title.x=element_blank()) +
  ylab("Accuracy")

ggsave(file="../graphs/accuracy_trialType.pdf",width=6,height=4.5)
```

# Redundant modifier trials:
```{r overinformativeness, echo=FALSE}
df = dd %>%
  filter(trialType == "redundantModification")

drows = nrow(df)
#table(df$clickedObjCondition)

df = df %>%
  mutate(targetFull=targetName) %>%
  separate(targetFull,c("targetobj","targetsize","targetcolor"))

#table(df$targetobj,df$correct)

toplot =  df %>%
  group_by(targetobj) %>%
  dplyr::summarise(Mean=mean(correct),CILow=ci.low(correct),CIHigh=ci.high(correct)) %>%
  ungroup() %>%
  mutate(YMin=Mean-CILow,YMax=Mean+CIHigh)

ggplot(toplot, aes(x=reorder(targetobj,Mean), y=Mean)) +
  geom_bar(position="dodge", stat="identity", fill="lightblue") +
  geom_errorbar(aes(ymin = YMin, ymax = YMax),width=.25, position=position_dodge(width = 0.9)) +
  scale_fill_manual(values=c("darkgreen","darkgoldenrod1")) +
  xlab("items") +
  ylab("Accuracy") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file="../graphs/accuracy_redmodifer.pdf",width=8,height=4)

```

**Was a color mentioned?**
```{r colormention, echo=FALSE}
colors = "blue|green|red|brown|black|clear|purple|pink|silver|gray|grey|white|yellow|violet|gold|silver|orange|rose|blu|purp|sky blu|greebn|yellwo|yello|silve|greeb|gren|yeelow"
df$colorMention = ifelse(grepl(colors, df$speakerMessages, ignore.case = TRUE), 1,0)
table(df$colorMention)
```

**Was size mentioned?**
```{r sizemention, echo=FALSE}
sizes = "big|small|biggest|smallest|tiny|huge|little|large|largest|tiniest|siz|biggesst|smallesst|smaller|bigger|smal"
df$sizeMention = ifelse(grepl(sizes, df$speakerMessages, ignore.case = TRUE), 1,0)
table(df$sizeMention)
```

**Was both a color and size mentioned?**
```{r colorsizesmention, echo=FALSE}
df$colorsizeMention = ifelse(df$colorMention==1&df$sizeMention==1,1,0)
table(df$colorsizeMention)
```

**Was an object name mentioned?**
```{r objectmention, echo=FALSE}
names  = "dress|apple|hair|pepper|glasses|ball|lighter|pot|button|clock|knife|tie|umbrella|spoon|shorts|tshirt|brush|box|skirt|socks|scissors|glove|lock|shoe|flower|bag|lipstick|toothbrush|car|baloon|balloon|chair|key|pants|cat|peper|paint brush|bell pepper|shirt|scissor|pan|shoes|backpack|necktie|sneaker|alarm clock|hand|sunglasses|paint brush|auto|paintbrush|knif";
df$objectMention = ifelse(grepl(names, df$speakerMessages, ignore.case = TRUE), 1, 0)
table(df$objectMention)
```

**In how many trials did the listener choose the right object?**
```{r fcorrect_selection, echo=FALSE}
table(df$correct)
non_target_click = drows-sum(df$correct)

print(paste("percentage of trials where target wasn't selected: ", non_target_click*100/drows)) #8.89
print("distribution accross conditions: ")
table(df$clickedObjCondition,df$correct)
```

**Exclude trials where target wasn't selected**
```{r clicktarget, echo=FALSE}
targets = df %>% filter(correct == "1")
nrow(targets) # 1133 trials

targets = targets %>%
  mutate(condition=clickedObjCondition) %>%
  mutate(UtteranceType = ifelse(colorMention==1 & sizeMention==1,"color and size",ifelse(colorMention==1,"only color",ifelse(sizeMention==1,"only size","other")))) %>%
  filter(UtteranceType!="other") %>%
  mutate(rowNum = row_number()) %>%
  mutate(onlyColorMention = ifelse(UtteranceType =="only color",1,0)) %>%
  mutate(onlySizeMention = ifelse(UtteranceType =="only size",1,0)) %>%
  mutate(RedundantProperty = ifelse(condition=="color_sufficient","size", ifelse(condition=="size_sufficient","color", NA))) %>%
  mutate(SufficientProperty = ifelse(condition=="color_sufficient","color", ifelse(condition=="size_sufficient","size", NA))) %>%
  mutate(RedExp3Cat = ifelse(UtteranceType=="color and size","redundant",ifelse((SufficientProperty=="color"&UtteranceType=="only color"),"minimal",ifelse((SufficientProperty=="size"&UtteranceType=="only size"),"minimal","underinformative")))) %>%
  mutate(RedExp2Cat = ifelse(RedExp3Cat=="redundant","redundant","not_redundant")) %>%
  mutate(redundant = ifelse(RedExp2Cat=="redundant",1,0)) %>%
  mutate(condition=ifelse(condition=="size_sufficient","size \n sufficient",ifelse(condition=="color_sufficient","color \n sufficient",NA))) %>%
  select(gameid,gameid,roundNum,condition,targetName,correct,speakerMessages,colorMention,sizeMention,UtteranceType,onlyColorMention,onlySizeMention,colorsizeMention,RedundantProperty,SufficientProperty,RedExp3Cat,RedExp2Cat,redundant)
```

**Reading corrected dataset for typos and mention of other modifiers**
TODO: manual correction + code 'other' modifiers
```{r manually_corrected, eval=FALSE, include=FALSE}
#tinfo = d %>% mutate(id = row_number()) %>% select(id,trialType)
manual = read.xlsx("../data/manual_correction.xlsx",1)
d = d %>% dplyr::mutate(id = row_number())
tomerge = manual %>% select(id,needCorrection,otherModifier)
d = merge(d, tomerge, by="id", all.x=TRUE, all.y=TRUE) 
d$otherMod = ifelse(is.na(d$otherModifier),"0","1")

shade  = "dark|light|shade|lite|sky|violet|bright|pale";
d$shade = ifelse(grepl(shade, d$otherModifier, ignore.case = TRUE), "1", "0")
size = "big|small|long|short|lengthy|tall|large";
d$size = ifelse(grepl(size, d$otherModifier, ignore.case = TRUE), "1", "0")
shape = "rectengular|rectengle|square";
d$shape = ifelse(grepl(shape, d$otherModifier, ignore.case = TRUE), "1", "0")

print(paste("percentage of trials where other modifiers were used: ", sum(as.numeric(d$otherMod))*100/drows)) #39.18
```

**Proportion of utterance by utterance type and trial type**
```{r mainplot, echo=FALSE}
agr = targets %>%
  gather(MentionType,mention,onlyColorMention:colorsizeMention) %>%
  group_by(RedundantProperty,condition,MentionType) %>%
  summarise(Mean=mean(mention),CIlow=ci.low(mention),CIhigh=ci.high(mention)) %>%
  ungroup() %>%
  mutate(Ymin=Mean-CIlow, Ymax=Mean+CIhigh) %>%
  mutate(MentionType=fct_recode(MentionType,"color and size"="colorsizeMention","only color"="onlyColorMention","only size"="onlySizeMention"))

ggplot(agr, aes(x=condition,y=Mean,fill=RedundantProperty)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = Ymin, ymax = Ymax),width=.25) +
  xlab("Trial type") +
  facet_wrap(~MentionType) + 
  scale_fill_manual(values=c(palette[1],palette[3])) +
  theme(legend.position = "none",axis.title.x = element_blank(),axis.text.x = element_text(size=7)) +
  ylab("Proportion of utterance")

ggsave(file="../graphs/proportion_overall.pdf",width=4,height=2)
```

```{r simple_model, echo=FALSE}
library(lme4)
library(languageR)
#centering condition 
tomodel = targets %>%
  mutate(condition = as.factor(condition)) %>%
  mutate(ccondition = as.numeric(condition) - mean(as.numeric(condition)))

m = glmer(redundant ~ ccondition + (1+ccondition|gameid) + (1|targetName), data=tomodel, family="binomial")
summary(m)
```

**Proportion of utterance type**
```{r utterancetype, echo=FALSE}
agr = targets %>%
  mutate(ThreeRedundant=ifelse(RedExp3Cat=="redundant",1,0)) %>%
  mutate(ThreeMinimal=ifelse(RedExp3Cat=="minimal",1,0)) %>%
  mutate(ThreeUnderinformative=ifelse(RedExp3Cat=="underinformative",1,0)) %>%
  gather(ThreeType,value,ThreeRedundant:ThreeUnderinformative) %>%
  group_by(RedundantProperty,ThreeType,condition) %>%
  summarise(Mean=mean(value),CIlow=ci.low(value),CIhigh=ci.high(value)) %>%
  ungroup() %>%
  mutate(Ymin=Mean-CIlow, Ymax=Mean+CIhigh) %>%
  mutate(ThreeType=fct_recode(ThreeType,"redundant"="ThreeRedundant","underinformative"="ThreeUnderinformative","minimal"="ThreeMinimal"))

ggplot(agr, aes(x=condition,y=Mean,fill=RedundantProperty)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = Ymin, ymax = Ymax),width=.25) +
  xlab("Trial type") +
  facet_grid(~ThreeType) + 
  scale_fill_manual(values=c(palette[1],palette[3])) +
  theme(legend.position = "none",axis.title.x = element_blank(),axis.text.x = element_text(size=7)) +
  ylab("Proportion of utterance")

ggsave(file="../graphs/proportion_threetype.png",width=4,height=2)
``` 

**Was negation used**   
Per attempt (not per trial)
```{r negation, echo=FALSE}
neg = "no|not"
df$negation = ifelse(grepl(neg, df$speakerMessages, ignore.case = TRUE), 1,0)

table(df$negation)
```

**Speaker messages that include negation**   
```{r negation2, echo=FALSE}
negation = df %>%
  filter(negation==1)

#negation$speakerMessages
```

**Order of unique color, noun, size (remove all other constituents)**
```{r order_3, echo=FALSE}
df_long = df %>%
  mutate(messages=speakerMessages) %>%
  separate(messages,c("attempt1","attempt2","attempt3","attempt4","attempt5"), sep="___") %>%
  gather(attempt,gloss,attempt1:attempt5) %>%
  drop_na(gloss) %>%
  select(gameid,speakerMessages,attempt,gloss)

df_long$replacedGloss = tolower(df_long$gloss)

pattern_rm = "blank|unique|different|load|unique|and|odd|nothing|empty|with|of"
df_long$remove=ifelse(str_detect(df_long$replacedGloss, pattern_rm),1,0)
df_long = df_long %>% filter(remove==0)
  
df_long$replacedGloss = gsub(colors,"color", df_long$replacedGloss)
df_long$replacedGloss = gsub(sizes,"size", df_long$replacedGloss)
df_long$replacedGloss = gsub(names,"noun", df_long$replacedGloss)

# should one, item and object be counted as noun
noun_like = "item|one|object"
df_long$replacedGloss = gsub(noun_like,"noun", df_long$replacedGloss)

toremove = "the|target|is|it's|very|it|!"
df_long$replacedGloss = gsub(toremove,"", df_long$replacedGloss)

df_long = df_long[!(df_long$replacedGloss==""), ]
df_long$replacedGloss = gsub("  "," ", df_long$replacedGloss)
df_long$replacedGloss = gsub("  "," ", df_long$replacedGloss)

# plot proportion not count
order = df_long %>%
  mutate(replacedGloss = trimws(replacedGloss,which=c("left"))) %>%
  mutate(replacedGloss = trimws(replacedGloss,which=c("right"))) %>%
  group_by(replacedGloss) %>%
  mutate(grp = ifelse(replacedGloss=="size","no noun",ifelse(replacedGloss=="noun","no noun",ifelse(replacedGloss=="color","no noun",ifelse(replacedGloss=="color size","no noun",ifelse(replacedGloss=="size color", "no noun", ifelse(replacedGloss=="color noun","pre-nominal",ifelse(replacedGloss=="size noun","pre-nominal",ifelse(replacedGloss=="color size noun","pre-nominal",ifelse(replacedGloss=="size color noun","pre-nominal",ifelse(replacedGloss=="noun color","post-nominal",ifelse(replacedGloss=="noun size","post-nominal",ifelse(replacedGloss=="noun size color","post-nominal",ifelse(replacedGloss=="noun color size","post-nominal",ifelse(replacedGloss=="color noun size","split",ifelse(replacedGloss=="size noun color","split","NA")))))))))))))))) %>%
  filter(grp != "NA") %>%
  mutate(only_color = ifelse(replacedGloss =="color",1,ifelse(replacedGloss =="color noun",1,ifelse(replacedGloss =="noun color",1,0)))) %>%
  mutate(only_size = ifelse(replacedGloss =="size",1,ifelse(replacedGloss =="size noun",1,ifelse(replacedGloss =="noun size",1,0)))) %>%
  mutate(only_noun = ifelse(replacedGloss =="noun",1,0)) %>%
  mutate(color_size = ifelse(replacedGloss =="color size",1,ifelse(replacedGloss =="size color",1,ifelse(replacedGloss =="color size noun",1,ifelse(replacedGloss =="color noun size",1,ifelse(replacedGloss =="size color noun",1,ifelse(replacedGloss =="size noun color",1,ifelse(replacedGloss =="noun color size",1,ifelse(replacedGloss =="noun size color",1,0))))))))) %>%
  mutate(preNominal = ifelse(grp=="pre-nominal",1,0)) %>%
  mutate(postNominal = ifelse(grp=="post-nominal",1,0)) %>%
  mutate(split = ifelse(grp=="split",1,0)) %>%
  mutate(noOrder = ifelse(grp=="no noun",1,0))
  
order_1 = order %>%
  gather(modType,value,preNominal:noOrder) %>%
  group_by(modType) %>%
  summarize(Mean=mean(value)) %>%
  mutate(modType = fct_recode(modType,"pre-nominal"="preNominal","post-nominal"="postNominal","no noun"="noOrder"))

order_1$modType_f = factor(as.factor(order_1$modType),levels=c("pre-nominal","post-nominal","split","no noun"))

ggplot(order_1, aes(x=modType_f,y=Mean, fill=modType)) +
  geom_bar(position="dodge", stat = "identity") +
  theme(axis.title.y=element_blank()) +
  scale_fill_manual(values= wes_palette("Darjeeling2")) +
  ylab("proportion") +
  xlab("modification type")+
  theme(legend.position = "none")

ggsave(file="../graphs/modification_type_english.pdf",width=4,height=3)

###########
order_2= order %>%
  filter(split==0) %>%
  gather(orderType,value,only_color:color_size) %>%
  group_by(grp,orderType) %>%
  summarize(Mean=mean(value)) %>%
  filter(Mean>0)
  
order_2$grp_f = factor(as.factor(order_2$grp),levels=c("pre-nominal","post-nominal","no noun"))

plt = wes_palette("Darjeeling2")

ggplot(order_2, aes(x=orderType,y=Mean,fill=grp_f)) +
  geom_bar(position="dodge", stat = "identity") +
  coord_flip() +
  scale_fill_manual(values= c(plt[3],plt[2],plt[1],plt[2])) +
  facet_wrap(~grp_f, scales = "free_y", ncol=1) +
  theme(axis.title.y=element_blank()) +
  ylab("proportion of utterance") +
  theme(legend.position = "none")
  
ggsave(file="../graphs/all_orders_english.pdf",width=5,height=4)
```

# Modifier ordering trials:
```{r adjordering, echo=FALSE}
da = dd %>%
  filter(trialType == "modifierOrdering") %>%
  mutate(targetName2=targetName) %>%
  separate(targetName2,c("targetobj","targetadj1","targetadj2"))

#table(da$targetobj,da$correct)

toplot =  da %>%
  group_by(targetobj) %>%
  dplyr::summarise(Mean=mean(correct),CILow=ci.low(correct),CIHigh=ci.high(correct)) %>%
  ungroup() %>%
  mutate(YMin=Mean-CILow,YMax=Mean+CIHigh)

ggplot(toplot, aes(x=reorder(targetobj,Mean), y=Mean)) +
  geom_bar(position="dodge", stat="identity", fill="lightblue") +
  geom_errorbar(aes(ymin = YMin, ymax = YMax),width=.25, position=position_dodge(width = 0.9)) +
  scale_fill_manual(values=c("darkgreen","darkgoldenrod1")) +
  xlab("items") +
  ylab("Accuracy")

ggsave(file="../graphs/accuracy_adjordering.pdf",width=8,height=4)
```

**Best objects: earring, pencil**  
```{r echo=FALSE, comment='', results='asis'}
da_earring = da %>%
  filter(targetobj=="earring") %>%
  select(targetName,speakerMessages)
knitr::kable(da_earring, floating.environment="sidewaystable")

da_pencil = da %>%
  filter(targetobj=="pencil") %>%
  select(targetName,speakerMessages)
knitr::kable(da_pencil, floating.environment="sidewaystable")
```

**Worst objects: spoon, hair**   
```{r echo=FALSE, comment='', results='asis'}
da_spoon = da %>%
  filter(targetobj=="spoon") %>%
  select(targetName,speakerMessages)
knitr::kable(da_spoon, floating.environment="sidewaystable")

da_hair = da %>%
  filter(targetobj=="hair") %>%
  select(targetName,speakerMessages)
knitr::kable(da_hair, floating.environment="sidewaystable")
```

